---
title: "Bakery Analysis"
author: "Rica Rebusit"
date: "2023-10-29"
output:
  html_document:
    df_print: paged
---

I will be using this dataset from Kaggle -> ![Click Me!]("https://www.kaggle.com/datasets/matthieugimbert/french-bakery-daily-sales/") to analyze and forecast sales of a french bakery from 2021 to 2022 based on:

date: Date order
time: Time order
ticket number: Identifier for every single transaction
article: Name of the product sold
quantity: quantity sold
unit_price: price per product (In Euros)

There are 234,005 observations in the dataset as well.

# Loading Necessary Libraries and Importing Data

```{r, warning = FALSE}
library(tidyverse)
library(gridExtra)
```

```{r}
bakery <- read.csv("Bakery sales.csv")
```

# EDA and Data Cleaning

```{r}
head(bakery)
```
```{r}
str(bakery)
```

```{r}
summary(bakery)
```


Checking for missing values
```{r}
sum(is.na(bakery))
#No missing values
```

Checking for duplicated rows
```{r}
sum(duplicated(bakery))
#No duplicated rows
```

Cleaning up variable names
```{r}
bakery <- select(bakery, -("X")) %>% #Dropping unnecessary column
  rename(quantity = Quantity) #Rename column Quantity with a lowercase q to keep variable uniformity
head(bakery)
```

Convert *unit_price* to numeric for analysis and sale forecasting. First getting rid of Euro symbol and replacing comma with a period
```{r}
bakery$unit_price <- gsub(" €", " ", bakery$unit_price) #Replaces Euro with empty space
bakery$unit_price <- gsub(",", ".", bakery$unit_price) #Replaces comma with period
head(bakery)
```

Now we can convert *unit_price* to numeric
```{r}
bakery$unit_price <- as.numeric(bakery$unit_price)
head(bakery)
```

Calculating total of each purchase
```{r}
bakery <- mutate(bakery, total = unit_price * quantity) #Mutate creates a new column
head(bakery)
```

For analysis, creating columns that separate date into year, month, day, and day name
```{r}
bakery$date <- as.Date(bakery$date)
bakery <- mutate(bakery, 
    year = format(bakery$date, "%Y"),
    month = format(bakery$date, "%m"),
    day = format(bakery$date, "%d"),
    day_name = format(bakery$date, "%A"))
head(bakery)
```

Checking for outliers
```{r}
grid.arrange(
  ggplot(bakery, aes(quantity)) + geom_boxplot(),
  ggplot(bakery, aes(total)) + geom_boxplot(),
  ncol = 2 
) 
```

It seems there is a negative value for qunatity and total, needs to be dropped
```{r}
bakery <- filter(bakery, !(quantity <= 0 | total <= 0))
head(bakery)
```

Since the dataset provides years 2021 and 2022, let's look at total in both years
```{r}
ggplot(bakery, aes(year, total, fill = year)) + geom_bar(stat = "Identity") + theme_minimal() + scale_fill_brewer(palette = "Accent") + labs(title = "Total Revenue (Year)") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
```

In months
```{r}
ggplot(bakery, aes(month, total, fill = month)) + geom_bar(stat = "Identity") + theme_minimal() + labs(title = "Total Revenue (Month)") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
```

In days
```{r}
ggplot(bakery, aes(day_name, total, fill = day_name)) + geom_bar(stat = "Identity") + theme_minimal() + labs(title = "Total Revenue (Day)") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
```

Comparing total revenue of both years
```{r}
ggplot(bakery, aes(month, total, fill = month)) + geom_bar(stat = "Identity") + facet_wrap(~ year) + labs(title = "Total Revenue (2021 vs 2022)") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
```

Looking at products in the bakery
```{r}
unique(bakery$article)
```

Which are the top 10 products?
```{r}
top10 <- bakery %>%
  group_by(article) %>%
  summarise(total = sum(total)) %>%
  arrange(desc(total)) %>%
  head(10)
top10
```

Looking at total from top 10
```{r}
ggplot(top10, aes(article, total, fill = article)) + geom_bar(stat = "Identity") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + labs(title = "Total (Top 10)") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
```

Which are the bottom 10 products?
```{r}
bottom10 <- bakery %>%
  group_by(article) %>%
  summarise(total = sum(total)) %>%
  arrange(total) %>%
  head(10)
bottom10
```
Looking at bottom 10 products
```{r}
ggplot(bottom10, aes(article, total, fill = article)) + geom_bar(stat = "Identity") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + labs(title = "Total (Bottom 10)") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
```

What is the revenue growth?

```{r}
bakery_jan_sept <- filter(bakery, month %in% c("01","02","03","04","05","06","07","08","09")) #Choosing months January-September because 2022 only goes up to September
bakery_jan_sept <- filter(bakery_jan_sept, year == "2021")
rev2021 <- sum(bakery_jan_sept[bakery_jan_sept$year == "2021", "total"], na.rm = TRUE)
rev2021
```

2022 revenue
```{r}
bakery2022 <- filter(bakery, year == "2022")
rev2022 <- sum(bakery2022[bakery2022$year == "2022", "total"], na.rm = TRUE)
rev2022
```

Calculating revenue growth
```{r}
revenue_growth <- ((rev2022 - rev2021) / rev2021) * 100
revenue_growth
```

Comparing the revenue growth from 2021 and 2022 of months January-September there is a 4.34% revenue growth. There is a possible assumption that the revenue growth would be higher if we had sales for October-December in 2022

# Forecasting Sales

Predicting future sales with quantity and total as the dependent variables

### Train and Test Data

It is a large dataset so it's best to do 80-20 split
```{r}
tra.tes <- sample(nrow(bakery), 0.8*nrow(bakery))
train <- bakery[tra.tes, ]
test <- bakery[-tra.tes, ]
```

### Linear Regression - Model Building

```{r}
bakery
```


To predict future sales, I will be using linear regression.
```{r}
set.seed(123)
#First with quantity as dependent variable
lm.quantity <- lm(quantity ~ article + unit_price + total + year + month + day_name, data = train)

#Now with total as dependent variable
lm.total <- lm(total ~ article + unit_price + quantity + year + month + day_name, data = train)
summary(lm.total)
```

# Linear Regression - Evaluation

```{r}
summary(lm.quantity)
```
The R-squared value is 75.03% the model fairly learned the relationships well

```{r}
summary(lm.total)
```

The R-squared is 88.02% which also seems to learn the relationships well. Better than the first model.


